Stack Safety as a Security Property

### 2020-07-07 main.tex
another pass on intro

But what {\em is} this ``stack safety'' of which we speak? Precise formal
definitions are lacking in the literature; instead, the notion of safety
is usually defined negatively, by enumerating the bad things that a safe system
prevents, such as corruption of return addresses, buffer overflows, or
use of uninitialized variables.
%with the notion usually being
%defined negatively: through examples of stack-based exploits
%\rb{usually though not always; as noted below there are other
%  approaches}\bcp{Yes, we have to be very careful about this}.
%\leo{Examples at a high level}: such as -- does not overwrite return addresses, or ...
We are only aware of one, quite recent, attempt to frame a positive definition.
\citet{Skorstengaard+19} define stack safety as
the conjunction of two properties, {\em local state encapsulation} and {\em
  well-bracketed control flow}.
They formalize these properties by defining an abstract capability-based machine
that ``satisfies [stack safety] by construction.''
%
While this is an important step in the right direction, we believe
there is still room to craft a more abstract and less intensional
characterization of stack safety, so that we are able to answer
questions about other enforcement mechansims. That is the goal of this
paper.
\leo{I took a stab. I think it's still to dense, but captures the
  feeling of today's discussion I think}
\apt{still needs tuning}\bcp{Yes, I think we need to go into a
  little more detail on what their definition looks like in order to
  justify why an alternative definition might be valuable.  What
  questions does their presentation not answer? What uses does it not
  support? E.g., can it be enforced efficiently? (I guess so, by
  capabilities, though I think their capabilities are more powerful
  than what Cheri actually provides, while our policies are exactly
  what Nick needs.)  Why would it be good to look at stack safety
  through the lens of security?  (Because it is fundamentally a
  security mechanism?)}  \rb{BCP idea: contrast intrinsic
  vs. extrinsic property definition.  Remember to talk about spatial
  vs temporal safety}

Our main contribution in this paper is an alternative formal presentation of
local state encapsulation and well-bracketed control flow using the
conceptual and technical framework of language-based security.
%
The underlying notion behind local state encapsulation is the intention to
protect callers from callees: a callee should be not be able to access or
overwrite private data in the caller's stack frame. We formalize this
intuition as a security property with both confidentiality and integrity
components.
%
Intuitively, {\em stack confidentiality} promises that a callee's
observable behavior is not affected by memory outside its stack frame
(e.g. a caller's private data or uninitialized memory).  {\em Stack
  integrity} guarantees that, if a function call ever returns, the caller's
stack data at the point of return will be identical to what it was before
the call.  \bcp{We need to point out somewhere that we are only dealing with
stack safety, though we discuss (somewhere) how our property can be combined
with the heap safety property of [Azevedo et al]}\leo{Should this be
  here or in the assumptions as presented in the running example?}

@@

To provide evidence that our definitions capture the inituitive notion
of stack safety, we show that they rule out a range of exploitable bad
behaviors.\apt{``from the literature''??}  To assess whether they are
implementable, we compare them against the proposed capability-based
enforcement mechanism of \citet{Skorstengaard+19} \apt{we really
  should} and the existing tag-based enforcement mechanism of
\citet{DBLP:conf/sp/RoesslerD18}.\leo{Is this where we would rephrase this
  because we would only talk about it in related work?}

In summary, we offer the following contributions:
\begin{itemize}
\item
  % \leo{More? Extensional}\bcp{We might need to invent a word for   this---''quasi-extensional''?}
  Abstract formal properties capturing the
  intuitive concepts of local state encapsulation and well-bracketed
  control flow~(\cref{sec:lse-and-wbcf}) in a generic setting. These properties
  rule out common examples of stack memory abuse.
\item
  A case study of the Depth Isolation policy of \citet{DBLP:conf/sp/RoesslerD18}, based on hardware tags~(\cref{sec:enforcement}).
  This policy enforces our stack safety property, and also a (provably) stronger property that is easier to test.
  \apt{If we have no testing, do we even want to talk about the testing
    property?} \bcp{Not here; just in future work.  I guess we need to see
    whether we wind up with any results at all about testing.}
\item
  A refinement of the case study that considers the more efficient Lazy Tagging and Clearing policy
  of \citet{DBLP:conf/sp/RoesslerD18}~(\cref{sec:lazy}).  This policy does not enforce our property, but we give a
  variant of the policy that does enforce a (provably) weaker variant of our
  property.
%
  \apt{Awkward, but the truth I fear.} \bcp{I agree the phrasing is awkward,
    but the point we can make is still strong: To make the property easier
    to enforce, we can slightly weaken its guarantees; the weaker version is
    a bit more complicated to state but still retains the pragmatic
    protections of the original.}\apt{To clarify: Nick's LT\&C policy uses
    depth-based tags, which do not enforce even our lazy policy. So we need to
    introduce a version of LT\&C that uses unique activation tags.}

>>>>

But what {\em is} this ``stack safety'' of which we speak? Precise formal
definitions are lacking in the literature; instead, the notion of safety
is usually defined negatively, by enumerating the bad things that a safe system
prevents, such as corruption of return addresses, buffer overflows, or
use of uninitialized variables.
%with the notion usually being
%defined negatively: through examples of stack-based exploits
%\rb{usually though not always; as noted below there are other
%  approaches}\bcp{Yes, we have to be very careful about this}.
%\leo{Examples at a high level}: such as -- does not overwrite return addresses, or ...
We are only aware of one, quite recent, attempt to frame a positive definition.
\citet{Skorstengaard+19} define stack safety as
the conjunction of two properties, {\em local state encapsulation} and {\em
  well-bracketed control flow}.
They formalize these properties by defining an abstract capability-based machine
that ``satisfies [stack safety] by construction.''
%
While this is an important step in the right direction, we believe
there is still room to craft a more abstract, generic characterization
of stack safety that is independent of any particular enforcement
mechanism.  In other words, we seek formal criteria that can be applied
against \emph{any} \apt{stack-based?} system to determine whether or not
the system is stack-safe.  
\apt{another stab.}
\leo{I took a stab. I think it's still to dense, but captures the
  feeling of today's discussion I think}
\apt{still needs tuning}\bcp{Yes, I think we need to go into a
  little more detail on what their definition looks like in order to
  justify why an alternative definition might be valuable.  What
  questions does their presentation not answer? What uses does it not
  support? E.g., can it be enforced efficiently? (I guess so, by
  capabilities, though I think their capabilities are more powerful
  than what Cheri actually provides, while our policies are exactly
  what Nick needs.)  Why would it be good to look at stack safety
  through the lens of security?  (Because it is fundamentally a
  security mechanism?)}  \rb{BCP idea: contrast intrinsic
  vs. extrinsic property definition.  Remember to talk about spatial
  vs temporal safety}

The main contribution of this paper is a novel formal presentation of
local state encapsulation and well-bracketed control flow using the
conceptual and technical framework of language-based security.
%
The underlying notion behind local state encapsulation is the intention to
protect callers from callees: a callee should be not be able to access or
overwrite private data in the caller's stack frame. We formalize this
intuition as a security property with both confidentiality and integrity
components.
%
Intuitively, {\em stack confidentiality} promises that a callee's
observable behavior is not affected by memory outside its stack frame
(e.g. a caller's private data or uninitialized memory).  {\em Stack
  integrity} guarantees that, if a function call ever returns, the caller's
stack data at the point of return will be identical to what it was before
the call.  \bcp{We need to point out somewhere that we are only dealing with
stack safety, though we discuss (somewhere) how our property can be combined
with the heap safety property of [Azevedo et al]}\leo{Should this be
  here or in the assumptions as presented in the running example?}

@@

To provide evidence that our definitions capture the inituitive notion
of stack safety, we show that they rule out a range of exploitable bad
behaviors.\apt{``from the literature''??}  To assess whether they are
implementable, we compare them against 
%proposed capability-based
%enforcement mechanism of \citet{Skorstengaard+19} \apt{we really
%  should} and
the existing tag-based enforcement mechanism of
\citet{DBLP:conf/sp/RoesslerD18}.
%\leo{Is this where we would rephrase this because we would only talk about it in related work?}

In summary, we offer the following contributions:
\begin{itemize}
\item
  % \leo{More? Extensional}\bcp{We might need to invent a word for   this---''quasi-extensional''?}
  Abstract formal properties capturing the
  intuitive concepts of strict local state encapsulation and well-bracketed
  control flow~(\cref{sec:lse-and-wbcf}) in a generic setting. These properties
  rule out common examples of stack memory abuse.
\item
  A case study of the Depth Isolation policy of \citet{DBLP:conf/sp/RoesslerD18}, based on hardware tags~(\cref{sec:enforcement}).
  This policy enforces our strict stack safety property, and also a (provably) stronger property that is easier to test.
  \apt{If we have no testing, do we even want to talk about the testing
    property?} \bcp{Not here; just in future work.  I guess we need to see
    whether we wind up with any results at all about testing.}
\item
  A refinement of the case study that considers the more efficient Lazy Tagging and Clearing policy
  of \citet{DBLP:conf/sp/RoesslerD18}~(\cref{sec:lazy}).  This policy does not enforce our strict
  local state encapsulation property,
  but a modified version of it using unique per-activation tags enforces 
  a (provably) weaker lazy variant of the property. 
%
% \apt{Awkward, but the truth I fear.} \bcp{I agree the phrasing is awkward,
%   but the point we can make is still strong: To make the property easier
%   to enforce, we can slightly weaken its guarantees; the weaker version is
%   a bit more complicated to state but still retains the pragmatic
%%   protections of the original.}\apt{To clarify: Nick's LT\&C policy uses
%   depth-based tags, which do not enforce even our lazy policy. So we need to
%   introduce a version of LT\&C that uses unique activation tags.}
